---
title: "ANN"
output: pdf_document
date: "2024-07-30"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Installing required packages

Here, I am installing the necessary packages and loading them.

```{r}
install.packages(c('neuralnet', 'kera','tensorflow'), dependencies = T)
install.packages("tidyverse")
library(neuralnet)
library(tidyverse)
```

## Data and processing

Here we get the data set for the model. We are converting the iris data set column into a factor. This is important for the neural network.

```{r}
iris<-iris%>%mutate_if(is.character, as.factor)
iris_sample <- iris[sample(nrow(iris), 10), ]
summary(iris)
```

## Splitting the data

Now we split the data into training and testing.
We will calculate the the number of rows that represent 80% of the dataset. Then randomly select the indices that will be used for training. Then, split the data.

Finally, we print the head for the test and train datasets.

```{r}
set.seed(254)
data_rows<-floor(0.80*nrow(iris))
data_rows

train_indices<-sample(c(1:nrow(iris)), data_rows)
train_indices

train_data<-iris[train_indices, ]
head(train_data)

test_data<-iris[-train_indices, ]
head(test_data)
```

## ANN model

Now we create a neural network model that will predict the species based on the features in the dataset.

Once we create the model, we then plot it

```{r}
model<-neuralnet(Species ~ Sepal.Length + Sepal.Width + Petal.Length + Petal.Width, data = train_data, hidden = c(4), linear.output = FALSE)
plot(model, rep='best')
```

## Evaluation

Now that the model has been created, we need to test and evaluate it, and see the accuracy

```{r}
pred<-predict(model, test_data)
pred
labels<-c("setosa", "versicolor", "virginica")
labels
prediction_label <- data.frame(max.col(pred)) %>% mutate(pred=labels[max.col.pred.]) %>% select(2) %>% unlist()
table(test_data$Species, prediction_label)

check = as.numeric(test_data$Species) == max.col(pred)
accuracy = (sum(check)/nrow(test_data))*100
print(accuracy)
```

## Layers

Now to see the effect the layers have. Currently, the model has one hidden layer with 4 neurons. But let's see different layers and how if affects the accuracy.

```{r}
model<-neuralnet(Species ~ Sepal.Length + Sepal.Width + Petal.Length + Petal.Width, data = train_data, hidden = c(2, 4), linear.output = FALSE)
plot(model, rep='best')

pred<-predict(model, test_data)
labels<-c("setosa", "versicolor", "virginica")
prediction_label <- data.frame(max.col(pred)) %>% mutate(pred=labels[max.col.pred.]) %>% select(2) %>% unlist()
table(test_data$Species, prediction_label)

check = as.numeric(test_data$Species) == max.col(pred)
accuracy = (sum(check)/nrow(test_data))*100
print(accuracy)
```

The above neural network has two hidden layers, with 2 and 4 neurons respectively

```{r}
model<-neuralnet(Species ~ Sepal.Length + Sepal.Width + Petal.Length + Petal.Width, data = train_data, hidden = c(2, 4, 4), linear.output = FALSE)
plot(model, rep='best')

pred<-predict(model, test_data)
labels<-c("setosa", "versicolor", "virginica")
prediction_label <- data.frame(max.col(pred)) %>% mutate(pred=labels[max.col.pred.]) %>% select(2) %>% unlist()
table(test_data$Species, prediction_label)

check = as.numeric(test_data$Species) == max.col(pred)
accuracy = (sum(check)/nrow(test_data))*100
print(accuracy)
```

The above neural network has three hidden layers, with 2, 4 and 4 neurons respectively

```{r}
model<-neuralnet(Species ~ Sepal.Length + Sepal.Width + Petal.Length + Petal.Width, data = train_data, hidden = c(2, 4, 4, 5), linear.output = FALSE)
plot(model, rep='best')

pred<-predict(model, test_data)
labels<-c("setosa", "versicolor", "virginica")
prediction_label <- data.frame(max.col(pred)) %>% mutate(pred=labels[max.col.pred.]) %>% select(2) %>% unlist()
table(test_data$Species, prediction_label)

check = as.numeric(test_data$Species) == max.col(pred)
accuracy = (sum(check)/nrow(test_data))*100
print(accuracy)
```

The above neural network has three hidden layers, with 2, 4, 4 and 5 neurons respectively

## Conclusion

Increasing the layers increases the accuracy for sure, but it also is unnecessary with this dataset. It is not that complex.